Automatic Matching of Bullet Lands

### 2015-12-22 imaging-paper.Rnw
Alicia's color
sweave options out
hopefully last style change

\documentclass[11pt]{article}
\usepackage[margin = 1in]{geometry}

\usepackage{lineno,hyperref}
\modulolinenumbers[5]
\usepackage{setspace}
\doublespacing

\usepackage[USenglish]{babel}
\usepackage[protrusion=true,expansion=true]{microtype} % Better typography
\usepackage{amsmath,amsfonts,amsthm} % Math packages
\usepackage[svgnames]{xcolor} % Enabling colors by their 'svgnames'
\usepackage[hang, small,labelfont=bf,up,textfont=it,up]{caption} % Custom captions under/above floats in tables or figures
\usepackage{booktabs} % Horizontal rules in tables
\usepackage{fix-cm}	 % Custom font sizes - used for the initial letter in the document
\usepackage{natbib}
\usepackage{float}
\usepackage{subcaption}

% declare the path(s) where your graphic files are
\graphicspath{{Figure/}{Images/}}
% and their extensions so you won't have to specify these with
% every instance of \includegraphics
\DeclareGraphicsExtensions{.pdf,.jpg,.png}

\usepackage{sectsty} % Enables custom section titles
\allsectionsfont{\usefont{OT1}{phv}{b}{n}} % Change the font of all section commands

\usepackage{fancyhdr} % Needed to define custom headers/footers
\pagestyle{fancy} % Enables the custom headers/footers
\usepackage{lastpage} % Used to determine the number of pages in the document (for "Page X of Total")

\usepackage{url}

\usepackage{color}
\newcommand{\hh}[1]{{\color{magenta} #1}}
\newcommand{\eh}[1]{{\color{blue} #1}}

    \renewcommand{\topfraction}{0.99}	% max fraction of floats at top
    \renewcommand{\bottomfraction}{0.99}	% max fraction of floats at bottom
% parameters for text pages
\setcounter{topnumber}{2}
    \setcounter{bottomnumber}{2}
    \setcounter{totalnumber}{4}     % 2 may work better
    \setcounter{dbltopnumber}{2}    % for 2-column pages
    \renewcommand{\dbltopfraction}{0.9}	% fit big float above 2-col. text
    \renewcommand{\textfraction}{0.07}	% allow minimal text w. figs
    %   Parameters for FLOAT pages (not text pages):
    \renewcommand{\floatpagefraction}{0.7}	% require fuller float pages
	% N.B.: floatpagefraction MUST be less than topfraction !!
    \renewcommand{\dblfloatpagefraction}{0.7}	% require fuller float pages

%----------------------------------------------------------------------------------------

\title{Random Forests for Automatic Matching Bullet-Lands}
\author{Eric Hare{\footnote{Corresponding author: erichare@iastate.edu}}, Heike Hofmann, Alicia Carriquiry\\
Department of Statistics and Statistical Laboratory, Iowa State University}

\begin{document}
\maketitle

\begin{abstract}
some change, just for fun
In 2009, the National Academy of Sciences published a report questioning the scientific validity of many forensic methods including firearm examination. Firearm examination is a forensic tool used to determine whether two bullets were fired from the same gun barrel. During the firing process, rifling, manufacturing defects, and impurities in the barrel create striation marks on the bullet. Identifying these striation markings in an attempt to match two bullets is the primary goal. We propose a framework for the analysis of the 3D surface measurements which transcribes the markings into a 2D plotting framework. This makes identification of matches easier and more rooted in science for both the expert and the lay person. The automatic matching routine we propose \hh{manages to (a) correctly identify lands with too much damage to be suitable for matching, and (b) correctly identify all 10,384 land-to-land matches.}
\end{abstract}

\paragraph{Keywords:} \hh{XXX not the right keywords yet} 3d Topological Surface Measurement, Data visualization, Machine Learning, Feature importance.

\newpage
\linenumbers

\tableofcontents

@@

\begin{figure}[hbtp]
  \centering
\begin{subfigure}[t]{.49\linewidth}\centering
\caption{View from the side.\label{fig:sidex3p}}{%
      \includegraphics[width=\linewidth]{images/sidex3p.png}
    }
\end{subfigure}    
\begin{subfigure}[t]{.49\linewidth}\centering
    \caption{View from the top.\label{fig:topx3p}}{%
    \includegraphics[width=\linewidth]{images/topx3p.png}
    }
\end{subfigure}
\caption{An example of a scan of a bullet land from one groove to the next. }
\end{figure}
Each fired bullet is provided in form of a set of six x3p files, where each file is a surface scan between adjacent grooves on the bullet, called a ``land". For notational simplicity, we refer to a particular land of some bullet as bullet X-Y, where X is the bullet identifier, and Y is the land number. An example of plotting one of these lands is given in Figures~\ref{fig:sidex3p} and~\ref{fig:topx3p}. These figures show side and top profiles of the land respectively. The tilt of the lines to the left in Figure~\ref{fig:topx3p} is not an artifact, but a direct and expected consequence of the spin induced by the rifling during the firing process. Depending on whether a barrel is rifled clockwise or counter-clockwise, the striations have a left or right tilt. The direction of the rifling is a class characteristic, or a feature that pertains to a particular class of firearms and is not unique at the individual bullet level.

@@

\begin{figure}[hbtp]
  \centering
<<fixedX, dependson='data', echo=FALSE, warning=FALSE, message=FALSE, fig.height=2, fig.width=6, out.width='0.5\\linewidth'>>=

@@

\begin{figure}[hbtp]
  \centering
\begin{subfigure}[b]{.49\linewidth}\centering
\caption{\label{fig:residuala}Residual structure at $x = 1.5625$ (bottom of the bullet).}
<<residual2, dependson='fixedX', echo=FALSE, warning=FALSE, fig.height=3, out.width='\\linewidth'>>=
qplot(data=subset(dbr111, x <= 1.5625), y, value-ypred, #colour=factor(x),
      geom="line", size=I(1)) +
#  scale_colour_brewer("x", palette="Paired") + 
  theme_bw() + 
  geom_hline(yintercept = 0, colour="grey50") +
  ylab(expression(paste("Residuals (in ",mu,"m)", sep=""))) + 
  theme(legend.position="bottom")
@
\end{subfigure}    
\begin{subfigure}[b]{.49\linewidth}\centering
\caption{\label{fig:residualb} Residual structure at $x = 100.00$}
<<residual, dependson='fixedX', echo=FALSE, warning=FALSE, fig.height=3, out.width='\\linewidth'>>=
#qplot(data=subset(dbr111, x <= 80*1.5625^2 & x >=75*1.5625^2), y, value-ypred,
qplot(data=subset(dbr111, x == 100), y, value-ypred,  #colour=factor(x), 
      geom="line", size=I(1)) +
#  scale_colour_brewer("x", palette="Paired") + 
  geom_hline(yintercept = 0, colour="grey50") +
  theme_bw() + 
  ylab(expression(paste("Residuals (in ",mu,"m)", sep=""))) + 
  theme(legend.position="bottom")
@
\end{subfigure}
\caption{\label{fig:residual} Residual structure of circular fits at two different cross sections. Both residual plots show systematic structures, indicating that a circular fit is not entirely appropriate.}
\end{figure}

@@

\begin{figure}[hbtp]
  \centering
<<circlefits, echo=FALSE, fig.width=10, fig.height=5, out.width='\\linewidth', warning=FALSE>>=
db2$land <- db2$part
db1$land <- db1$part
qplot(y, resid, data=subset(db2, x == 100), #colour=factor(x), 
      geom="line", size=I(.75), colour=I("grey70")) + 
  facet_wrap(~land, ncol=3, labeller="label_both") + 
  scale_colour_brewer(palette="Paired") +
  theme_bw() + theme(legend.position="bottom") + 
  geom_line(aes(y, resid, group = x), colour="black", size=.75, alpha=0.5,
             data = filter(db1, land==5, x == 100)[,c("y", "resid", "x")]) +
  ylab("Residuals from circular fit") +
  ggtitle("Bullet 1-5 in black")
@
\caption{Circle fit for each cross section of bullet 2, with bullet 1-5 overlaid.\label{fig:circlefits} Bullet 1-5 matches best with bullet 2-1.}
\end{figure}

@@

\begin{figure}[hbtp]
  \centering
\begin{subfigure}[t]{\linewidth}\centering
\caption{\label{fig:loess_step1}Step 1 the loess fit algorithm. For a fixed cross section ($x$ held constant at 243.75 $\mu m$) of surface measurements for bullet 1-5 are plotted across different values of $y$.}
<<loess_step1, echo=FALSE, fig.width=10, fig.height=3.5, out.width='.5\\linewidth', warning=FALSE,  message=FALSE>>=
br111_bullet <- get_crosscut(paste(datadir,"Br1 Bullet 1-5.x3p", sep = "/"))
br121_bullet <- get_crosscut(paste(datadir,"Br1 Bullet 2-1.x3p", sep = "/"))
br122_bullet <- get_crosscut(paste(datadir,"Br1 Bullet 2-2.x3p", sep = "/"))
br123_bullet <- get_crosscut(paste(datadir,"Br1 Bullet 2-3.x3p", sep = "/"))
br124_bullet <- get_crosscut(paste(datadir,"Br1 Bullet 2-4.x3p", sep = "/"))
br125_bullet <- get_crosscut(paste(datadir,"Br1 Bullet 2-5.x3p", sep = "/"))
br126_bullet <- get_crosscut(paste(datadir,"Br1 Bullet 2-6.x3p", sep = "/"))

p1 <- qplot(y, value, data = br111_bullet, size=I(1)) + theme_bw() + ylab("Surface measurements")
p1
@
\end{subfigure}
\begin{subfigure}[t]{\linewidth}\centering
\caption{\label{fig:loess_step3}Step 2 of the loess fit algorithm. The  height values are smoothed twice with a smoothing factor of 35. The orange rectangle shows an example of the smoothing window. Grooves and peaks are detected, if they are not within the same window.}
<<loess_step3, echo=FALSE, fig.width=10, fig.height=3.5, out.width='.5\\linewidth', warning=FALSE,  message=FALSE>>=
subdata <- br111_bullet
smoothfactor <- 35

@@

p3
@
\end{subfigure}
\begin{subfigure}[t]{\linewidth}\centering
\caption{\label{fig:loess_step47}Steps 3 -- 6 of the loess fitting algorithm. After smoothing the $y$ values left and right minima are detected (marked by vertical lines, red indicating peaks and blue indicating valleys). Values outside these boundaries are removed (shown in grey)}
<<loess_step47, echo=FALSE, fig.width=10, fig.height=3.5, out.width='.5\\linewidth', warning=FALSE,  message=FALSE>>=
    lengthdiff <- length(subdata$value) - length(smoothed_truefalse)

@@

\begin{figure}[hbtp]
  \centering
\begin{subfigure}[t]{.49\linewidth}\centering
\caption{\label{fig:loess_fit} Loess fit for bullet 1-5.}
<<loess_fit, echo=FALSE, fig.width=10, fig.height=3.5, out.width='\\linewidth', warning=FALSE,  message=FALSE>>=

@@

my.loess$fitted + ylab("Surface measurements")
@
\end{subfigure}
\begin{subfigure}[t]{.49\linewidth}\centering
\caption{\label{fig:loess_resid} Residuals of loess fit for bullet 1-5.}
<<loess_resid, echo=FALSE, fig.width=10, fig.height=3.5, out.width='\\linewidth', warning=FALSE>>=
my.loess$resid + ylab("Residuals of loess Fit")
@
\end{subfigure}
\caption{Fit and residuals of a loess fit to bullet 1-5 (Barrel 1).}
\end{figure}

\begin{figure}[hbtp]
  \centering
<<loessfits, echo=FALSE, fig.width=10, fig.height=5, out.width='\\linewidth', warning=FALSE>>=
br121.groove <- get_grooves(br121_bullet)
br122.groove <- get_grooves(br122_bullet)
br123.groove <- get_grooves(br123_bullet)
br124.groove <- get_grooves(br124_bullet)
br125.groove <- get_grooves(br125_bullet)
br126.groove <- get_grooves(br126_bullet)

@@

lines <- matches %>% group_by(lineid) %>% summarise(
  meany = mean(y, na.rm=T),
  miny = min(y, na.rm=T),
  maxy = max(y, na.rm=T) + 1.5625,
  match = isMatch(lineid, type),
  type = type1[1]
)
lines <- subset(lines, lineid != 0)
@
\begin{figure}[hbtp]
\centering
\begin{subfigure}[b]{.49\linewidth}\centering
\caption{Cross section at $x = 100$.\label{fig:crosscut}}{%
<<crosscuts, echo=FALSE, dependson='twolands100', fig.width=7, fig.height=3, out.width='\\textwidth'>>=
qplot(y, resid, group=bullet, colour=bullet, data=lof, 
      geom="line") + 
  scale_colour_manual("", values=cols) + theme_bw() +
  theme(legend.position = c(.9, .2),
        legend.background = element_rect(fill=alpha('white', 0.4))) + 
  ylab("Residuals of loess fit")
@
    }
\end{subfigure}    
\begin{subfigure}[b]{.49\linewidth}\centering
    \caption{Adjusted (bullet 1-5 horizontally shifted) cross sections at $x = 100$.\label{fig:crosscutX}}{%
<<crosscutsX, echo=FALSE, dependson='twolands100', fig.width=7, fig.height=3, out.width='\\textwidth'>>=
# qplot(y, resid, group=bullet, colour=bullet, data=lofX, 
#       geom="line", alpha=I(0.6)) + 
#   scale_colour_brewer(palette="Set1") + theme_bw() +
#   theme(legend.position="none")
qplot(y+min(lof$y), resid, group=bullet, colour=bullet, data=lofY, 
      geom="line") + xlab("y") +
  scale_colour_manual("", values=cols) + theme_bw() +
  theme(legend.position="none") + ylab("Residuals of loess fit")
@
    }
\end{subfigure}
\caption{\label{fig:cross100}Lines showing surface cross sections at $x = 100$. A horizontal shift of the values of bullet 1-5 to the right shows the similarity of the striation marks.}
\end{figure}

@@

\begin{figure}[hbtp]
  \centering
<<ccf, echo=FALSE, fig.width = 7, fig.height = 3, out.width = '.5\\linewidth'>>=
lag <- ccf$lag[which.max(ccf$acf)]
qplot(x=ccf$lag, xend=ccf$lag, y = 0, yend= ccf$acf, geom="segment", colour = I("grey50")) + 
  theme_bw() + ylab("Correlation") + xlab("Lag k") +
  geom_segment(aes(x = lag, xend = lag, y = 0, yend =  max(ccf$acf)), colour="black") + 
  scale_x_continuous(breaks = c(-100, -50, lag, 0, 50, 100),
                     minor_breaks = c(-75,-25,25, 75))
@
\caption{\label{fig:ccf}Cross-correlation between the two cross sections shown in Figure~\ref{fig:crosscut} at lags between -100 and 100. At a lag of -10 the correlation peaks, indicating the largest amount of agreement between the cross sections. Figure~\ref{fig:crosscutX} shows the lag-shifted cross sections.}
\end{figure}

\subsection{Varying cross sections}
The cross section at which a comparison is done matters -- as cross sections are further apart, more pronounced differences between the cross sections show up. This poses both a caveat to matching attempts as well as an opportunity \hh{for quality control}: on the one hand we have to be aware of which cross section level was used in a matching, but on the other hand we can use multiple cross sections of the same bullet for an initial assessment of its quality. Both of these situations are assessed and discussed in more detail below: 
Figure~\ref{fig:crosscuts} shows a sequence of cross sections for bullet 1-5 (barrel 1) that are taken 50$\mu m$ apart, between 150$\mu m$  and 400$\mu m$. These are compared to a cross section at 100$\mu m$. Initially this comparison constitutes an almost perfect match between the two cross sections. However, the match quickly deteriorates with increasing distance to non-matches at a cross section of $x = 300\mu m$ or higher.  Only if the cross section is within 150$\mu m$ do we get a good visual match even when we know that the same bullet surface is being used. 
Given that we have to expect some variation in nominally the same cross section levels due to manual alignments in microscopes, we should take cross section levels into account in the automatic matching routine by evaluating matches at several cross sections. 
\begin{figure}[hbtp]
  \centering
<<crosscuts-vary, echo=FALSE, fig.width = 12, fig.height = 7, out.width = '\\linewidth', cache=TRUE, warning = FALSE>>=
paths <- file.path(datadir, dir(datadir))
crosscuts <- seq(100, 400, by = 50)
lof <- processBullets(paths[5], x = crosscuts)
lof$bullet <- paste(lof$bullet, lof$x)

@@

\begin{figure}[hbtp]
  \centering
<<crosscuts-vary-b31, echo=FALSE, fig.width = 12, fig.height = 7, out.width = '\\linewidth', cache=TRUE, warning = FALSE>>=
paths <- file.path(datadir, dir(datadir))
crosscuts <- seq(100,250, by=25)
lof <- processBullets(paths[37], x = crosscuts)
lof$bullet <- paste(lof$bullet, lof$x)

@@

\begin{figure}[hbtp]
\centering
    \begin{subfigure}[t]{\linewidth}\centering
\caption{Loess smooth at $x = 100$ (span is 0.03).\label{fig:smooth}}{%
<<smooth, echo=FALSE, dependson='twolands100', fig.width=7, fig.height=3.5, out.width='.5\\textwidth', warning=FALSE>>=
qplot(data=lofX, y, resid, colour=I("grey30"), size=I(0.75), geom="line", group=bullet) +
  geom_line(aes(y=l30), colour="grey70", size=0.5) +
  facet_grid(bullet~.) +
#  scale_colour_manual("", values=cols) +
  theme_bw() + ylab("") + 
  theme(legend.position="none")
@
    }
\end{subfigure}
\begin{subfigure}[t]{\linewidth}\centering
\caption{Using a rolling median peaks and grooves are identified for each cross section. \label{fig:smoothcutb}}{%
<<smoothcut, echo=FALSE, dependson='twolands100', fig.width=7, fig.height=3.5, out.width='.5\\textwidth', warning=FALSE>>=
#images <- file.path(datadir, dir(datadir))
  lof <- processBullets(paths = images[c(5,7)], x = 100)
  lof <- bulletSmooth(lof)
  bAlign = bulletAlign(lof)
  lofX <- bAlign$bullet

@@

\begin{figure}[hbtp]
  \centering
<<barrel2, echo=FALSE, fig.width = 10, fig.height=6, out.width='\\linewidth', warning=FALSE>>=

>>>>

% Template for the submission to:
%   The Annals of Probability           [aop]
%   The Annals of Applied Probability   [aap]
%   The Annals of Statistics            [aos] 
%   The Annals of Applied Statistics    [aoas]
%   Stochastic Systems                  [ssy]
%
%Author: In this template, the places where you need to add information
%        (or delete line) are indicated by {???}.  Mostly the information
%        required is obvious, but some explanations are given in lines starting
%Author:
%All other lines should be ignored.  After editing, there should be
%no instances of ??? after this line.

% use option [preprint] to remove info line at bottom
% journal options: aop,aap,aos,aoas,ssy
% natbib option: authoryear
\documentclass[aoas]{imsart}

\usepackage{amsthm,amsmath,natbib}
\RequirePackage[colorlinks,citecolor=blue,urlcolor=blue]{hyperref}

% provide arXiv number if available:
\arxiv{arXiv:0000.0000}

% put your definitions there:
\startlocaldefs
\usepackage[hang, small,labelfont=bf,up,textfont=it,up]{caption} % Custom captions under/above floats in tables or figures
\usepackage{subcaption}
\usepackage{lineno,hyperref}
\modulolinenumbers[5]

\newcommand{\hh}[1]{{\color{magenta} #1}}
\newcommand{\eh}[1]{{\color{blue} #1}}
\newcommand{\ac}[1]{{\color{darkgreen} #1}}

\endlocaldefs

\begin{document}

\begin{frontmatter}

% "Title of the paper"
\title{Automatic Matching of Bullet Lands}
\runtitle{Automatic Matching of Bullet Lands}

% indicate corresponding author with \corref{}
\begin{aug}
\author{\fnms{Eric} \snm{Hare}\corref{}\ead[label=e1]{erichare@iastate.edu}\thanksref{t1, m1}},
\author{\fnms{Heike} \snm{Hofmann}\ead[label=e2]{hofmann@iastate.edu}\thanksref{m1}}
\and
\author{\fnms{Alicia} \snm{Carriquiry}\ead[label=e3]{alicia@iastate.edu}\thanksref{m1}}

\thankstext{t1}{Thanks to somebody \hh{XXX acknowledge the grant here? XXX}} 
\affiliation{Iowa State University\thanksmark{m1}}

\address{Address of the Authors\\
Department of Statistics and Statistical Laboratory\\ Snedecor Hall\\
Ames Iowa 50011-1210\\
\printead{e1}\\
\phantom{E-mail:\ }\printead*{e2}\\
\phantom{E-mail:\ }\printead*{e3}}

\runauthor{E. Hare et al.}
\end{aug}
\begin{abstract}
In 2009, the National Academy of Sciences published a report questioning the scientific validity of many forensic methods including firearm examination. Firearm examination is a forensic tool used to determine whether two bullets were fired from the same gun barrel. During the firing process, rifling, manufacturing defects, and impurities in the barrel create striation marks on the bullet. Identifying these striation markings in an attempt to match two bullets is the primary goal. We propose a framework for the analysis of the 3D surface measurements which transcribes the markings into a 2D plotting framework. This makes identification of matches easier and more rooted in science for both the expert and the lay person. The automatic matching routine we propose \hh{manages to (a) correctly identify lands with too much damage to be suitable for matching, and (b) correctly identify all 10,384 land-to-land matches.}
\end{abstract}

%\begin{keyword}[class=MSC]
%\kwd[Primary ]{}
%\kwd{}
%\kwd[; secondary ]{}
%\end{keyword}

\begin{keyword}
\kwd{3d Topological Surface Measurement}
\kwd{Data visualization}
\kwd{Machine Learning}
\kwd{Feature importance}
\kwd{Cross-correlation function}
\end{keyword}

\end{frontmatter}

\linenumbers

@@

\begin{figure}[hbtp]
  \centering
\begin{subfigure}[t]{.49\textwidth}\centering
\caption{View from the side.\label{fig:sidex3p}}{%
      \includegraphics[width=\textwidth]{images/sidex3p.png}
    }
\end{subfigure}    
\begin{subfigure}[t]{.49\textwidth}\centering
    \caption{View from the top.\label{fig:topx3p}}{%
    \includegraphics[width=\textwidth]{images/topx3p.png}
    }
\end{subfigure}
\caption{An example of a scan of a bullet land from one groove to the next. }
\end{figure}
Each fired bullet is provided in form of a set of six x3p files, where each file is a surface scan between adjacent grooves on the bullet, called a ``land". For notational simplicity, we refer to a particular land of some bullet as bullet X-Y, where X is the bullet identifier, and Y is the land number. An example of plotting one of these lands is given in Figures~\ref{fig:sidex3p} and~\ref{fig:topx3p}. These figures show side and top profiles of the land respectively. The tilt of the lines to the left in Figure~\ref{fig:topx3p} is not an artifact, but a direct and expected consequence of the spin induced by the rifling during the firing process. Depending on whether a barrel is rifled clockwise or counter-clockwise, the striations have a left or right tilt. The direction of the rifling is a class characteristic, or a feature that pertains to a particular class of firearms and is not unique at the individual bullet level.

@@

\begin{figure}[hbtp]
  \centering
<<fixedX, dependson='data', echo=FALSE, warning=FALSE, message=FALSE, fig.height=2, fig.width=6, out.width='0.5\\textwidth'>>=

@@

\begin{figure}[hbtp]
  \centering
\begin{subfigure}[b]{.49\textwidth}\centering
\caption{\label{fig:residuala}Residual structure at $x = 1.5625$ (bottom of the bullet).}
<<residual2, dependson='fixedX', echo=FALSE, warning=FALSE, fig.height=3, out.width='\\textwidth'>>=
qplot(data=subset(dbr111, x <= 1.5625), y, value-ypred, #colour=factor(x),
      geom="line", size=I(1)) +
#  scale_colour_brewer("x", palette="Paired") + 
  theme_bw() + 
  geom_hline(yintercept = 0, colour="grey50") +
  ylab(expression(paste("Residuals (in ",mu,"m)", sep=""))) + 
  theme(legend.position="bottom")
@
\end{subfigure}    
\begin{subfigure}[b]{.49\textwidth}\centering
\caption{\label{fig:residualb} Residual structure at $x = 100.00$}
<<residual, dependson='fixedX', echo=FALSE, warning=FALSE, fig.height=3, out.width='\\textwidth'>>=
#qplot(data=subset(dbr111, x <= 80*1.5625^2 & x >=75*1.5625^2), y, value-ypred,
qplot(data=subset(dbr111, x == 100), y, value-ypred,  #colour=factor(x), 
      geom="line", size=I(1)) +
#  scale_colour_brewer("x", palette="Paired") + 
  geom_hline(yintercept = 0, colour="grey50") +
  theme_bw() + 
  ylab(expression(paste("Residuals (in ",mu,"m)", sep=""))) + 
  theme(legend.position="bottom")
@
\end{subfigure}
\caption{\label{fig:residual} Residual structure of circular fits at two different cross sections. Both residual plots show systematic structures, indicating that a circular fit is not entirely appropriate.}
\end{figure}

@@

\begin{figure}[hbtp]
  \centering
<<circlefits, echo=FALSE, fig.width=10, fig.height=5, out.width='\\textwidth', warning=FALSE>>=
db2$land <- db2$part
db1$land <- db1$part
qplot(y, resid, data=subset(db2, x == 100), #colour=factor(x), 
      geom="line", size=I(.75), colour=I("grey70")) + 
  facet_wrap(~land, ncol=3, labeller="label_both") + 
  scale_colour_brewer(palette="Paired") +
  theme_bw() + theme(legend.position="bottom") + 
  geom_line(aes(y, resid, group = x), colour="black", size=.75, alpha=0.5,
             data = filter(db1, land==5, x == 100)[,c("y", "resid", "x")]) +
  ylab("Residuals from circular fit") +
  ggtitle("Bullet 1-5 in black")
@
\caption{Circle fit for each cross section of bullet 2, with bullet 1-5 overlaid.\label{fig:circlefits} Bullet 1-5 matches best with bullet 2-1.}
\end{figure}

@@

\begin{figure}[hbtp]
  \centering
\begin{subfigure}[t]{\textwidth}\centering
\caption{\label{fig:loess_step1}Step 1 the loess fit algorithm. For a fixed cross section ($x$ held constant at 243.75 $\mu m$) of surface measurements for bullet 1-5 are plotted across different values of $y$.}
<<loess_step1, echo=FALSE, fig.width=10, fig.height=3.5, out.width='.5\\textwidth', warning=FALSE,  message=FALSE>>=
br111_bullet <- get_crosscut(paste(datadir,"Br1 Bullet 1-5.x3p", sep = "/"))
br121_bullet <- get_crosscut(paste(datadir,"Br1 Bullet 2-1.x3p", sep = "/"))
br122_bullet <- get_crosscut(paste(datadir,"Br1 Bullet 2-2.x3p", sep = "/"))
br123_bullet <- get_crosscut(paste(datadir,"Br1 Bullet 2-3.x3p", sep = "/"))
br124_bullet <- get_crosscut(paste(datadir,"Br1 Bullet 2-4.x3p", sep = "/"))
br125_bullet <- get_crosscut(paste(datadir,"Br1 Bullet 2-5.x3p", sep = "/"))
br126_bullet <- get_crosscut(paste(datadir,"Br1 Bullet 2-6.x3p", sep = "/"))

p1 <- qplot(y, value, data = br111_bullet, size=I(1)) + theme_bw() + ylab("Surface measurements")
p1
@
\end{subfigure}
\begin{subfigure}[t]{\textwidth}\centering
\caption{\label{fig:loess_step3}Step 2 of the loess fit algorithm. The  height values are smoothed twice with a smoothing factor of 35. The orange rectangle shows an example of the smoothing window. Grooves and peaks are detected, if they are not within the same window.}
<<loess_step3, echo=FALSE, fig.width=10, fig.height=3.5, out.width='.5\\textwidth', warning=FALSE,  message=FALSE>>=
subdata <- br111_bullet
smoothfactor <- 35

@@

p3
@
\end{subfigure}
\begin{subfigure}[t]{\textwidth}\centering
\caption{\label{fig:loess_step47}Steps 3 -- 6 of the loess fitting algorithm. After smoothing the $y$ values left and right minima are detected (marked by vertical lines, red indicating peaks and blue indicating valleys). Values outside these boundaries are removed (shown in grey)}
<<loess_step47, echo=FALSE, fig.width=10, fig.height=3.5, out.width='.5\\textwidth', warning=FALSE,  message=FALSE>>=
    lengthdiff <- length(subdata$value) - length(smoothed_truefalse)

@@

\begin{figure}[hbtp]
  \centering
\begin{subfigure}[t]{.49\textwidth}\centering
\caption{\label{fig:loess_fit} Loess fit for bullet 1-5.}
<<loess_fit, echo=FALSE, fig.width=10, fig.height=3.5, out.width='\\textwidth', warning=FALSE,  message=FALSE>>=

@@

my.loess$fitted + ylab("Surface measurements")
@
\end{subfigure}
\begin{subfigure}[t]{.49\textwidth}\centering
\caption{\label{fig:loess_resid} Residuals of loess fit for bullet 1-5.}
<<loess_resid, echo=FALSE, fig.width=10, fig.height=3.5, out.width='\\textwidth', warning=FALSE>>=
my.loess$resid + ylab("Residuals of loess Fit")
@
\end{subfigure}
\caption{Fit and residuals of a loess fit to bullet 1-5 (Barrel 1).}
\end{figure}

\begin{figure}[hbtp]
  \centering
<<loessfits, echo=FALSE, fig.width=10, fig.height=5, out.width='\\textwidth', warning=FALSE>>=
br121.groove <- get_grooves(br121_bullet)
br122.groove <- get_grooves(br122_bullet)
br123.groove <- get_grooves(br123_bullet)
br124.groove <- get_grooves(br124_bullet)
br125.groove <- get_grooves(br125_bullet)
br126.groove <- get_grooves(br126_bullet)

@@

lines <- matches %>% group_by(lineid) %>% summarise(
  meany = mean(y, na.rm=T),
  miny = min(y, na.rm=T),
  maxy = max(y, na.rm=T) + 1.5625,
  match = isMatch(lineid, type),
  type = type1[1]
)
lines <- subset(lines, lineid != 0)
@
\begin{figure}[hbtp]
\centering
\begin{subfigure}[b]{.49\textwidth}\centering
\caption{Cross section at $x = 100$.\label{fig:crosscut}}{%
<<crosscuts, echo=FALSE, dependson='twolands100', fig.width=7, fig.height=3, out.width='\\textwidth'>>=
qplot(y, resid, group=bullet, colour=bullet, data=lof, 
      geom="line") + 
  scale_colour_manual("", values=cols) + theme_bw() +
  theme(legend.position = c(.9, .2),
        legend.background = element_rect(fill=alpha('white', 0.4))) + 
  ylab("Residuals of loess fit")
@
    }
\end{subfigure}    
\begin{subfigure}[b]{.49\textwidth}\centering
    \caption{Adjusted (bullet 1-5 horizontally shifted) cross sections at $x = 100$.\label{fig:crosscutX}}{%
<<crosscutsX, echo=FALSE, dependson='twolands100', fig.width=7, fig.height=3, out.width='\\textwidth'>>=
# qplot(y, resid, group=bullet, colour=bullet, data=lofX, 
#       geom="line", alpha=I(0.6)) + 
#   scale_colour_brewer(palette="Set1") + theme_bw() +
#   theme(legend.position="none")
qplot(y+min(lof$y), resid, group=bullet, colour=bullet, data=lofY, 
      geom="line") + xlab("y") +
  scale_colour_manual("", values=cols) + theme_bw() +
  theme(legend.position="none") + ylab("Residuals of loess fit")
@
    }
\end{subfigure}
\caption{\label{fig:cross100}Lines showing surface cross sections at $x = 100$. A horizontal shift of the values of bullet 1-5 to the right shows the similarity of the striation marks.}
\end{figure}

@@

\begin{figure}[hbtp]
  \centering
<<ccf, echo=FALSE, fig.width = 7, fig.height = 3, out.width = '.5\\textwidth'>>=
lag <- ccf$lag[which.max(ccf$acf)]
qplot(x=ccf$lag, xend=ccf$lag, y = 0, yend= ccf$acf, geom="segment", colour = I("grey50")) + 
  theme_bw() + ylab("Correlation") + xlab("Lag k") +
  geom_segment(aes(x = lag, xend = lag, y = 0, yend =  max(ccf$acf)), colour="black") + 
  scale_x_continuous(breaks = c(-100, -50, lag, 0, 50, 100),
                     minor_breaks = c(-75,-25,25, 75))
@
\caption{\label{fig:ccf}Cross-correlation between the two cross sections shown in Figure~\ref{fig:crosscut} at lags between -100 and 100. At a lag of -10 the correlation peaks, indicating the largest amount of agreement between the cross sections. Figure~\ref{fig:crosscutX} shows the lag-shifted cross sections.}
\end{figure}

\subsection{Varying cross sections}
The cross section at which a comparison is done matters -- as cross sections are further apart, more pronounced differences between the cross sections show up. This poses both a caveat to matching attempts as well as an opportunity \hh{for quality control}: on the one hand we have to be aware of which cross section level was used in a matching, but on the other hand we can use multiple cross sections of the same bullet for an initial assessment of its quality. Both of these situations are assessed and discussed in more detail below: 
Figure~\ref{fig:crosscuts} shows a sequence of cross sections for bullet 1-5 (barrel 1) that are taken 50$\mu m$ apart, between 150$\mu m$  and 400$\mu m$. These are compared to a cross section at 100$\mu m$. Initially this comparison constitutes an almost perfect match between the two cross sections. However, the match quickly deteriorates with increasing distance to non-matches at a cross section of $x = 300\mu m$ or higher.  Only if the cross section is within 150$\mu m$ do we get a good visual match even when we know that the same bullet surface is being used. 
Given that we have to expect some variation in nominally the same cross section levels due to manual alignments in microscopes, we should take cross section levels into account in the automatic matching routine by evaluating matches at several cross sections. 
\begin{figure}[hbtp]
  \centering
<<crosscuts-vary, echo=FALSE, fig.width = 12, fig.height = 7, out.width = '\\textwidth', cache=TRUE, warning = FALSE>>=
paths <- file.path(datadir, dir(datadir))
crosscuts <- seq(100, 400, by = 50)
lof <- processBullets(paths[5], x = crosscuts)
lof$bullet <- paste(lof$bullet, lof$x)

@@

\begin{figure}[hbtp]
  \centering
<<crosscuts-vary-b31, echo=FALSE, fig.width = 12, fig.height = 7, out.width = '\\textwidth', cache=TRUE, warning = FALSE>>=
paths <- file.path(datadir, dir(datadir))
crosscuts <- seq(100,250, by=25)
lof <- processBullets(paths[37], x = crosscuts)
lof$bullet <- paste(lof$bullet, lof$x)

@@

\begin{figure}[hbtp]
\centering
    \begin{subfigure}[t]{\textwidth}\centering
\caption{Loess smooth at $x = 100$ (span is 0.03).\label{fig:smooth}}{%
<<smooth, echo=FALSE, dependson='twolands100', fig.width=7, fig.height=3.5, out.width='.5\\textwidth', warning=FALSE>>=
qplot(data=lofX, y, resid, colour=I("grey30"), size=I(0.75), geom="line", group=bullet) +
  geom_line(aes(y=l30), colour="grey70", size=0.5) +
  facet_grid(bullet~.) +
#  scale_colour_manual("", values=cols) +
  theme_bw() + ylab("") + 
  theme(legend.position="none")
@
    }
\end{subfigure}
\begin{subfigure}[t]{\textwidth}\centering
\caption{Using a rolling median peaks and grooves are identified for each cross section. \label{fig:smoothcutb}}{%
<<smoothcut, echo=FALSE, dependson='twolands100', fig.width=7, fig.height=3.5, out.width='.5\\textwidth', warning=FALSE>>=
#images <- file.path(datadir, dir(datadir))
  lof <- processBullets(paths = images[c(5,7)], x = 100)
  lof <- bulletSmooth(lof)
  bAlign = bulletAlign(lof)
  lofX <- bAlign$bullet

@@

\begin{figure}[hbtp]
  \centering
<<barrel2, echo=FALSE, fig.width = 10, fig.height=6, out.width='\\textwidth', warning=FALSE>>=
